---
title: "WUP_04"
author: "Karen Gomez"
date: "Summer 2023"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)

```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

We first need to take a look at the data to attain the correct helper file. Then, we will upload the data and examine what results we are given. According to what we saw in ourdata, WORDSUM is a numerical variable and DEGREE is a factor variable. Which means we will need to upload the numerical~categorical helper file.  

```{r}
ourdata<- read.csv(file="GSScsv.csv")
```

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
basicCN <- ggplot(ourdata,aes(x=DEGREE,y=WORDSUM))

basicCN + geom_boxplot()
```

The graduate box shows a median that is much higher than the high school and lt high school box. Which means that in the WORDSUM test there is a great difference between graduate degrees and high school and lt high school degrees. 

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~DEGREE,data=ourdata)
```

The numerical results show that the average for graduate school is 10.1 and the average for junior college and bachelor degree is 9.9 and 9.2 respectively. Indicating that there is not much of a difference between graduate school, junior college, and bachelor degrees. However, there is a greater difference between graduate school and lt high school because the mean for graduate school is 10.1, whereas, the mean for lt high school is 7.8.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~DEGREE,data=ourdata)
summary.aov(model1)
```

The inferential results show a p-value of 0.000236. Which represents the chances that we could get the data under the same conditions, resulting that there is no difference between each degree. According to the data, there seems to be a difference. Therefore, we reject the null hypothesis. 

```{r}
#we specify the margins starting at the bottom and proceeding clockwise - thus 8.7 is the left side margin to hold the names of the factors
par(mar = c(3, 18.2, 2, 2))
plot(TukeyHSD(model1),las=2)
```

There are two conditions where the score on the degree are statistically different. Which are between high school and graduate school and the other is between lt high school and graduate school.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

In conclusion, we first looked at the problem we were given which was a numerical~categorical problem. It had 5 different categories and we ended up with 10 different t-test where we looked at the confidence intervals between the two means. The graduate degrees and the high school degree were greatly different and the graduate and lt high school degrees were also different.
